---
documentclass: article
fontsize: 10pt # Options Include 10pt, 11pt, or 12pt
header-includes:
- \usepackage{crimson}
- \usepackage[T1]{fontenc}
- \usepackage[margin=1in]{geometry}
- \renewcommand{\baselinestretch}{1}
- \usepackage{calc}
- \usepackage{enumitem}
- \usepackage{changepage}
output:
  pdf_document:
    number_sections: no # Yes if you wanted numbered sections
  word_document: default
urlcolor: blue
---

\vspace{-85pt}
\text{\bfseries\LARGE{MEMORANDUM}}
\newline
\vspace{-10pt}
\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}
\vspace{-18pt}
\setlist[2]{noitemsep}
\begin{description}[leftmargin=!,itemsep=-1ex,labelwidth=\widthof{Subject:  }]
  \item[To:] Santa Barbara Government 
  \item[From:] Sara Orofino and Jeremy Knox, Master of Environmental Science and Management Students
  \item[Date:] May 22, 2019
  \item[Subject:] \bfseries{Economics of a Vessel Speed Reduction Program in Santa Barbara Channel}
\end{description}
\vspace{-10pt}
\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}
\vspace{-25pt}

<!---THE MEMO BODY. EDITS ALL PLAIN TEXT--->

## BACKGROUND
\vspace{-10pt} 

The coast of California is home to a diverse marine ecosystem. Among the most important components are whales. Every year, over 20 whales are struck by container ships along the West Coast of the United States.[^1] The Santa Barbara Channel, off the coast of Calfiornia stretching from Point Conception to Point Mugu, is frequently visited by three endangered whale species: blue, fin and humpback. This is due to the unique location of the SB Channel where cold water from the north combines with warmer water from the south. Unfortunately, this channel is also a key shipping area for container ships. These ships transport 90% of the world's goods and travel through SB Channel more than 2,700 times every year[^2]. One proposed soltuion to reduce the frequency of whale strikes is the vessel speed reduction program (VSR). Reducing the speed of these ships would result in a 60% decrease in risk of whale strikes. The following analysis looks to quantify the economic benefits and costs of this VSR program. 


## METHODS
\vspace{-10pt} 

**Data**: A contingent valuation survey to 500 Santa Barbara County residents was given to determine the local demand for a VSR program. Variables measured: (1) _risk_: level of risk reduction, (2) _bid_: annual payment for the household, (3) _vote_: 1 is yes, 0 is no, (4) _NEP_ : measure of environmental concern, (5) _age_: categorical, (6) _income_: categorical.[^3]   

**Model**: Linear regression model of _vote_ on _age_, _income_, _NEP_, _bid_ and _risk_.[^4]



## ANALYSIS
\vspace{-10pt} 


To analyze Santa Barbara respondents probability of voting yes on a VSR program we use a linear regression model. Given that reducing the risk of whales strikes by 20% saves five whales, our model suggests the willingness to pay for saving one whale is 2.78(UNITS?), on average ceteris paribus.[^5] Furthermore, implementing a 60% reduction VSR program would result in an average willingness to pay of 71.48(UNITS?).[^6] Reducing the risk by this much would save 15 whales. Santa Barbara residents will recieve these benefits but the cost will be spread across the country (from a price increase on goods transported by these shipping vessels).The cost to implement a 60% reduction VSR is \$7 million. Santa Barbara residents would recieve a benefit of over \$10 billion giving a surplus of over $3.7 million.[^7]


Ships that reduce their speed not only reduce the risk of colliding with whales but also emit less carbon. An alternative to the price increase consumers of products aboard these ships face is a carbon credit program. There are carbon trading markets that could potentially offset the cost of implmenting a 60% reduction program. Our analysis shows that a carbon market that prices carbon at \$50 per ton would offset the cost of reducing speed thereby incentivising ships to reduce their speed.[^8] In this scenario, the benefit of 15 whales being saved to Santa Barbara Residents would be over \$10 billion while the cost would go to all tax payers and stakeholders who fund the carbon offeset program.  


## CONCLUSION
\vspace{-10pt} 

Given that the benefits outweigh the costs by 




[^1]: Redfern et al., 2013 cited from WhaleStrikes Group Porject

[^2]: Hakim, 2014 cited from WhaleStrikes Group Porject

[^3]: WhaleStrikes Group Porject

[^4]: See Technical Appendix: Linear model is used, instead of logistic, because the ease of interpretation. Significance of variables was not analyzed and averages/medians were used for calculations. 

[^5]: See Technical Appendix: (2)

[^6]: See Technical Appendix: (4)

[^7]: See Technical Appendix: (5) & (6)

[^8]: See Technical Appendix: (7)

\newpage



## Technical Appendix
\vspace{-10pt} 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}

# Packages
library(tidyverse)


# Read in data 
whales <- read_csv("HW3Data.csv")

#Create a version of the data with numbers for income and age to find the mode for use in part 4:  
whales_numeric <- whales %>%
  mutate(income_num = ifelse(whales$income == "very_rich", 4,
                             ifelse(whales$income == "rich", 3,
                                    ifelse(whales$income == "middle", 2,
                                           ifelse(whales$income == "poor", 1,
                                                  ifelse(whales$income == "one_percent", 5, 0)))))) %>% 
  mutate(age_num = ifelse(whales$age == "tothirty", 1,
                          ifelse(whales$age == "toforty", 2,
                                 ifelse(whales$age == "tofifty", 3,
                                        ifelse(whales$age == "tosixty", 4,
                                               ifelse(whales$age == "oversixty", 5,0))))))

# Find the median of income and age for use in part 4:
inc_median <- median(whales_numeric$income_num) #median income bracket is 3 = rich 
age_median <- median(whales_numeric$age_num) #median age bracket is 3 = tofifty


```

### 1. Linear Probability Model 

Create a linear probility model that predicts a respondent's probability of voting "yes" on the ballot based on their age, income, NEP score, the program's risk reduction, and cost of the program to that respondent.  

```{r, include=FALSE}
# Votes: 1 = "yes", 0 = "no"  

whales_lm <- lm(vote ~ age + income + NEP + bid + risk, data = whales)
summary(whales_lm)

int <- whales_lm$coefficients[1]
age_50 <- whales_lm$coefficients[2]
age_40 <- abs(whales_lm$coefficients[3])
age_60 <- abs(whales_lm$coefficients[4])
age_30 <- whales_lm$coefficients[5]
inc_one <- whales_lm$coefficients[6]
inc_poor <- whales_lm$coefficients[7]
inc_rich <- whales_lm$coefficients[8]
inc_veryrich <- whales_lm$coefficients[9]
nep <- whales_lm$coefficients[10]
bid <- abs(whales_lm$coefficients[11])
risk <- whales_lm$coefficients[12]
```
**Regression Model:**  
  
$Probability(Voting~Yes) = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) - `r round(age_40, digits=4)`(Age~to~40) + `r round(age_50, digits=4)`(Age~to~50) - `r round(age_60, digits=4)`(Age~to~60) + `r round(inc_one, digits=4)`(Income~One~Percent) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(inc_veryrich,digits=4)`(Income~Very~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk~Reduction)$

**Coefficient Interpretation:**  
*Age: Reference Level Over 65*    
  - to 30: All else being equal, a person in the age bracket of to 30 would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(age_30, digits=4)`$ higher than a person in the age bracket over 65.          
  - to 40: All else being equal, a person in the age bracket of to 40 would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(age_40, digits=4)`$ lower than a person in the age bracket over 65.      
  - to 50: All else being equal, a person in the age bracket of to 50 would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(age_50, digits=4)`$ higher than a person in the age bracket over 65.       
  - to 60: All else being equal, a person in the age bracket of to 60 would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(age_60, digits=4)`$ lower than a person in the age bracket over 65.    
    
*Income: Reference Level Middle*   
  - Poor: All else being equal, a person with an income level of one percent would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(inc_poor, digits=4)`$ higher than a person with a medium income level.     
  - Rich: All else being equal, a person with an income level of one percent would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(inc_rich, digits=4)`$ higher than a person with a medium income level.     
  - Very Rich: All else being equal, a person with an income level of one percent would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(inc_veryrich,digits=4)`$ higher than a person with a medium income level.   
  - One Percent: All else being equal, a person with an income level of one percent would be expected to have a probability of voting yes on a vessel speed reduction program that is, on average, $`r round(inc_one, digits=4)`$ higher than a person with a medium income level.   
    
*NEP:* For every 1 unit increase in environmental concern, we would expect the probability of voting yes on a vessel speed reduction program to increase by $`r round(nep, digits=4)`$, if all other variables are equal.
  
*Bid:*  For every 1 dollar increase in annual household payment, we expect the probability of voting yes on a vessel speed reduction program to decrease by $`r round(bid, digits=4)`$, if all other variables are equal.
  
*Risk:* For every 1 unit increase in risk reduction, we would expect the probability of voting yes on a vessel speed reduction program to increase by $`r round(risk, digits=4)`$, if all other variables are equal.    


### 2. Value of Prevented Whale Deaths  

Reducing the risk of whale strikes by 20% saves five whales every year. Based on this, the vessel speed reduction by 4% saves a single whale every year. To find the value of each individual whale saved find the willingess to pay for vessel speed reduction programs of 0% and compare to the willingness to pay for vessel speed reduction of 4%.     


**Risk Reduction 0%**  
Assume the probability of voting yes is the average of the votes ($p = 0.714$), assume an age to 30, income rich, and the average NEP (38.366), solve for the willingness to pay for the program using:

$0.714 = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(0)$


```{r, include=FALSE}
# Find the average of the yes votes:

avg_vote <- mean(whales$vote)

# Find the average of the NEP:

nep_avg <- mean(whales$NEP)

# Find the bid if risk reduction is zero:
bid_zero <- (int + age_30 + inc_rich + (nep*nep_avg) + (risk*0) - avg_vote)/bid

```

$Willingness~to~Pay = `r round(bid_zero, digits=2)`$    

**Risk Reduction 4%**  
Again, assume the probability of voting yes is the average of the votes ($p = 0.714$), assume an age to 30, income rich, and the average NEP (38.366), solve for the willingness to pay for the program using:  

$0.714 = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(4)$  


```{r, include=FALSE}

# Find bid if risk reduction is four:  

bid_four <- (int + age_30 + inc_rich + (nep*nep_avg) + (risk*4) - avg_vote)/bid

```

$Willingness~to~Pay = `r round(bid_four, digits=2)`$ 

The value of a single whale is the difference between the willingness to pay for a vessel speed reduction program at 4% and at 0%.  

```{r, include = FALSE}

value <- bid_four - bid_zero

```

$Individual~Whale~Value = `r round(value, digits=2)`$  


### 3. Estimated Willingness to Pay for a Vessel Speed Reduction Program  

#### a. Choose three participants at random
Using a random number generator select three participants:        
- 38   NEP:32  Income:Rich    Age:to30    
- 44   NEP:51  Income:Poor    Age:to40  
- 102  NEP:51  Income:Middle  Age:to60  

```{r, include=FALSE}
# Subset of data for 60% VSR program 
whale_60 <- whales %>%
  filter(risk == 60)

# Randomly select three integers between 1 and 106 without replacement 

sample(1:106, 3, replace = FALSE)

# Outcome - 38, 102, 44

# Create new dataframe with the random respondents information
vsr60_rand1 <- whale_60 %>% 
  slice(38,44,102)
 

```

#### b. Predict willingness to pay for 60% VSR program  

Assume the probability of voting yes the average of all the yes votes ($p = 0.714$), calculate the willingness to pay using the following equations:      

Individual 38:  
$`r avg_vote` = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

Individual 44:  
$`r avg_vote` = `r round(int, digits = 4)` - `r round(age_40, digits=4)`(Age~to~40) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$  

Individual 122:  
$`r avg_vote` = `r round(int, digits = 4)`- `r round(age_60, digits=4)`(Age~to~60) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid)+ `r round(risk, digits=4)`(60)$  


```{r, include = FALSE}

# Plug the values for randoms into the linear regression and solve for bid 

bid_38 <- (int+age_30+inc_rich+(nep*32)+(risk*60)-avg_vote)/bid

bid_44 <- (int+age_40+inc_poor+(nep*51)+(risk*60)-avg_vote)/bid

bid_122 <- (int+age_60+(nep*51)+(risk*60)-avg_vote)/bid
```

Individual 38:  
$Willingness~to~Pay = `r round(abs(bid_38), digits=2)`$    

Individual 44:  
$Willingness~to~Pay = `r round(abs(bid_44), digits=2)`$  

Individual 122:  
$Willingness~to~Pay = `r round(abs(bid_122), digits=2)`$  


### 4. Santa Barbara Estimated Willingness to Pay for VSR Program   

Again assume the probability of voting yes the average of voting yes ($p = 0.714$). Use the median income bracket (rich), the median age bracket (to 50), and average NEP (38.366) to calculate willingness to pay using: 

$`r avg_vote` = `r round(int, digits = 4)` + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(age_50, digits=4)`(Age~to~50) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$  


```{r, include = FALSE}
# Find the mode of the income and age to use in the equation:


# Use the average NEP, median income (rich), and middle age bracket (to 50) to find the average bid

bid_sb <- (int + age_50 + inc_rich + (nep*nep_avg) + (risk*60) - avg_vote)/bid

```

$Average~Santa~Barbara~County~Household~Willingness~to~Pay = `r round(abs(bid_sb), digits=2)`$   


### 5. Total benefits to Santa Barbara residents (population = 150,000). 

Risk = 60%  
Cost = 7 million 

Calculate Total Benefits using the following equation:  
Total Benefits = $`r round(bid_sb, digits=2)`~per~household * 150,000~households$  

```{r, include=FALSE}
total_benefit <- bid_sb * 150000

surplus <- total_benefit - 7000000
```


Total Benefits = $`r round(total_benefit, digits =2)`$  

### 6. Do Benefits Outweight Costs

Yes, based only on these benefits the benefits do outweight the costs.

Cost = 7 million  
Benefits = $`r total_benefit`$  

$Surplus = `r total_benefit` - 7000000 = `r surplus`$  


### 7. Price of Carbon Credits  

Assume that for any ship transiting the Santa Barbara Channel the speed reduction tha results in a 60% risk reduction costs the shipper 1000 but will result in 20 fewer tons of carbon dioxide per transit.   

Shipping companies will adopt the VSR for self interested reasons if the value of the carbon credits is equal to the cost.   

$20*Z = 1000$  
$Z = 50/ton$  


### 8. Whales Saved by Carbon Credits   
Assume the value of the carbon credit is 50/ton and all ships reduce speed to achieve the 60% risk reduction.      

Number of Whales Saved:  
20% Reduction = 5 Whales saved     
$Whales~Saved = 5*3 = 15$  

The social value of the whale reduction program is the total benefit of that program minus the cost. Since the carbon credits would avoid the 7 million cost of the program, the social value is just the total benefit of the program.     

$Social~Value = `r total_benefit`$  
