---
title: "ESM 204 Assignment 3"
author: "Jeremy Knox, Sara Orofino"
date: "5/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}

# Packages
library(tidyverse)


# Read in data 
whales <- read_csv("HW3Data.csv")
```

###1. Linear Probability Model 

Create a linear probility model that predicts a respondent's probability of voting "yes" on the ballot based on their age, income, NEP score, the program's risk reduction, and cost of the program to that respondent.  

```{r, include=FALSE}
# Votes: 1 = "yes", 0 = "no" - use binomial logistic regression 

whales_blr <- glm(vote ~ age + income + NEP + bid + risk, family = "binomial", data = whales)
summary(whales_blr)

int <- whales_blr$coefficients[1]
age_50 <- whales_blr$coefficients[2]
age_40 <- whales_blr$coefficients[3]
age_60 <- whales_blr$coefficients[4]
age_30 <- whales_blr$coefficients[5]
inc_one <- whales_blr$coefficients[6]
inc_poor <- whales_blr$coefficients[7]
inc_rich <- whales_blr$coefficients[8]
inc_veryrich <- whales_blr$coefficients[9]
nep <- whales_blr$coefficients[10]
bid <- whales_blr$coefficients[11]
risk <- whales_blr$coefficients[12]
```
Regression Model:  
$Logodds(Yes~Vote) = `r int` + `r age_30`(Age~to~30) + `r age_40`(Age~to~40) + `r age_50`(Age~to~50) + `r age_60`(Age~to~60) + `r inc_one`(Income~One~Percent) + `r inc_poor`(Income~Poor) + `r inc_rich`(Income~Rich) + `r inc_veryrich`(Income~Very~Rich) + `r nep`(NEP) + `r bid`(Bid) + `r risk`(Risk~Reduction)$
