---
title: "ESM 204 Assignment 3"
author: "Jeremy Knox, Sara Orofino"
date: "5/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}

# Packages
library(tidyverse)


# Read in data 
whales <- read_csv("HW3Data.csv")
```

###1. Linear Probability Model 

Create a linear probility model that predicts a respondent's probability of voting "yes" on the ballot based on their age, income, NEP score, the program's risk reduction, and cost of the program to that respondent.  

```{r, include=FALSE}
# Votes: 1 = "yes", 0 = "no" - use binomial logistic regression 

whales_lm <- lm(vote ~ age + income + NEP + bid + risk, data = whales)
summary(whales_lm)

int <- whales_lm$coefficients[1]
age_50 <- whales_lm$coefficients[2]
age_40 <- abs(whales_lm$coefficients[3])
age_60 <- abs(whales_lm$coefficients[4])
age_30 <- whales_lm$coefficients[5]
inc_one <- whales_lm$coefficients[6]
inc_poor <- whales_lm$coefficients[7]
inc_rich <- whales_lm$coefficients[8]
inc_veryrich <- whales_lm$coefficients[9]
nep <- whales_lm$coefficients[10]
bid <- abs(whales_lm$coefficients[11])
risk <- whales_lm$coefficients[12]
```
Regression Model:  
  
$Logodds(Voting~Yes) = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) - `r round(age_40, digits=4)`(Age~to~40) + `r round(age_50, digits=4)`(Age~to~50) - `r round(age_60, digits=4)`(Age~to~60) + `r round(inc_one, digits=4)`(Income~One~Percent) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(inc_veryrich,digits=4)`(Income~Very~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk~Reduction)$

**Coefficient Interpretation:**  
*Age:*    
  - to 30:   
  - to 40:   
  - to 50:   
  - to 60:   
*Income:*   
  - One Percent:  
  - Poor:  
  - Rich:  
  - Very Rich:  
*NEP:*    
*Bid:*    
*Risk:*    

###2. Value of Prevented Whale Deaths  


###3. Estimated Willingness to Pay for a Vessel Speed Reduction Program  

####a.Choose three participants at random
Using a random number generator select three participants      
- 38  
- 44 
- 102

```{r, include=FALSE}
# Subset of data for 60% VSR program 
whale_60 <- whales %>%
  filter(risk == 60)

# Randomly select three integers between 1 and 106 without replacement 

sample(1:106, 3, replace = FALSE)

# Outcome - 38, 102, 44

# Create new dataframe with the random respondents information
vsr60_rand1 <- whale_60 %>% 
  slice(38,44,102)
 

```

###b. Create a regression model to predict willingness to pay for 60% VSR program  

```{r, include = FALSE}
vsr60_lm <- lm(bid ~ NEP + income + age, data = whale_60)
summary(vsr60_lm)


int_60 <- vsr60_lm$coefficients[1]
nep_60 <- abs(vsr60_lm$coefficients[2])
inc_one_60 <- vsr60_lm$coefficients[3]
inc_poor_60 <- abs(vsr60_lm$coefficients[4])
inc_rich_60 <- vsr60_lm$coefficients[5]
inc_veryrich_60 <- vsr60_lm$coefficients[6]
age_50_60 <- abs(vsr60_lm$coefficients[7])
age_40_60 <- abs(vsr60_lm$coefficients[8])
age_60_60 <- abs(vsr60_lm$coefficients[9])
age_30_60 <- vsr60_lm$coefficients[10]

```

Model for estimating willingness to pay:  

$Bid = 